---
title: "Stan analysis 1 - Single subject"
author: "Kenta Yoshida"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: hide
---

# Setup and load data
```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(rstan)
library(ggmcmc)

rstan_options(auto_write=T)
options(mc.cores=parallel::detectCores())

data.pk   <- read_csv("../data/sim_pk_20170521.csv") 
data.subj <- read_csv("../data/subj_dose_20170521.csv") 
```

# Summary
Multiple ideas worked to avoid label switching, (1) weak prior on all parameters, (2) limit of Vd lower value (with weak priors), (3) ordered variables for KA and KEL, (4) limit ka with CL and VD, (5) provide weak priors and initial values.  
(5) to be picked for consistency with population model. 



# Data quick look

```{r}
data.pk %>%
  filter(ID==1) %>% 
  ggplot(aes(TIME, CONC)) +
  geom_line() +
  geom_point() 

full_join(data.pk, data.subj) %>% 
  ggplot(aes(TIME, CONC, group=ID)) +
  geom_line() +
  geom_point() +
  facet_wrap(~DOSE) +
  scale_y_log10()

```




# Strong prior for ka

```{r}
data.pk.id1 <- data.pk %>% filter(ID==1)

init <- function(){
    list(KA = exp(rnorm(1, log(0.5), 0.2)),
         CL = exp(rnorm(1, log(0.5), 0.2)),
         VD = exp(rnorm(1, log(5), 0.2)),
         s_Y = runif(1, 0.5, 2))
}

data <- 
  list(N    = nrow(data.pk.id1),
       TIME = data.pk.id1$TIME,
       DOSE = 10,
       Y    = data.pk.id1$CONC)


fit.stan <-
  stan(file = "mod_00_single_subj.stan", 
       data = data, init = init)

fit.stan
```


## Diagnostic plots

```{r}
# http://xavier-fim.net/packages/ggmcmc/#using-ggmcmc

fit.param <- ggs(fit.stan)
list.param <- c("KA", "CL", "VD", "s_Y")

fit.param.plot <- 
  fit.param %>% filter(Parameter %in% c(list.param))

ggs_density(fit.param.plot)
ggs_traceplot(fit.param.plot)
ggs_autocorrelation(fit.param.plot)

ggs_Rhat(fit.param.plot)
```


## Prediction vs Observed

```{r plot pred vs obs}
mcmc.sample <- rstan::extract(fit.stan)

y.pred.interval <- 
  mcmc.sample$y_new %>% 
  apply(MARGIN=2, quantile, prob=c(0.05, 0.5, 0.95)) %>% 
  t() %>% 
  tbl_df()

bind_cols(data.pk.id1,
          y.pred.interval) %>% 
  ggplot(aes(TIME, `50%`)) +
  geom_line() +
  geom_ribbon(aes(ymin=`5%`, ymax=`95%`),alpha=0.1) +
  geom_point(data=data.pk.id1, aes(TIME,CONC))

```




