---
title: "Stan analysis 2 - Population PK, with or without IIV"
author: "Kenta Yoshida"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: hide
---

# Setup and load data
```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(rstan)

rstan_options(auto_write=T)
options(mc.cores=parallel::detectCores())

data.pk   <- read_csv("../data/sim_pk_20170521.csv") 
data.subj <- read_csv("../data/subj_dose_20170521.csv")

data.pk.plot <- 
  mutate(data.pk, ID=factor(ID), DOSE_LEVEL=factor(DOSE_LEVEL))
```

# Summary
Strong prior worked for model w/o IIV, but with IIV, label switching issue persisted  
Tried ordered models but seemed like prior/limit settings are tricky, didn't really work

Best circumvention was to limit lower value of VD based on physiological rationale, both for typical VD and individual VD

The models in the blog are as follows:

* Without IIV, limit Vd
* With IIV, limit Vd

# Data quick look

```{r}

data.pk.plot %>% 
  ggplot(aes(TIME, CONC, group=ID, color=DOSE_LEVEL)) +
  geom_line() +
  geom_point() +
  facet_wrap(~DOSE_LEVEL) +
  scale_y_log10()

```


# Without IIV, strong prior

```{r}
data <- 
  list(N    = nrow(data.pk),
       N_ID = nrow(data.subj),
       ID   = data.pk$ID,
       TIME = data.pk$TIME,
       DOSE = data.subj$DOSE,
       Y    = data.pk$CONC)

fit.stan <-
  stan(file = "mod_01_noIIV_strong_prior_ka.stan", data = data)
```

## Prediction vs Observed

```{r plot pred vs obs}
mcmc.sample <- rstan::extract(fit.stan)

# Plot for each subject
y.pred.interval <- 
  mcmc.sample$y_new %>% 
  apply(MARGIN=2,quantile, prob=c(0.05, 0.5, 0.95)) %>% 
  t() %>% 
  tbl_df()

bind_cols(data.pk,
          y.pred.interval) %>% 
  filter(ID<=9)  %>% 
  mutate(ID=factor(ID)) %>% 
  ggplot(aes(TIME, `50%`, color=ID)) +
  facet_wrap(~ID) +
  geom_line() +
  geom_ribbon(aes(ymin=`5%`, ymax=`95%`, fill=ID),alpha=0.1) +
  geom_point(data=data.pk %>% filter(ID<=9) %>% mutate(ID=factor(ID)),
             aes(TIME,CONC)) +
  ylab("CONC")


# Plot for each dose level

dat.fit.plot.each.dose <- 
  ## Convert to tibble
  mcmc.sample$y_new %>% t() %>% tbl_df() %>% 
  ## Re-groupe predicted values according to TIME and DOSE_LEVEL
  bind_cols(data.pk,.) %>% 
  gather(key,prediction,-(ID:DOSE_LEVEL)) %>% 
  group_by(TIME,DOSE_LEVEL) %>%
  ## Calc quantiles
  summarize(`5%` =quantile(prediction, probs=0.05),
            `50%`=quantile(prediction, probs=0.5),
            `95%`=quantile(prediction, probs=0.95)) %>% 
  ungroup() %>% 
  mutate(DOSE_LEVEL=factor(DOSE_LEVEL))

ggplot(dat.fit.plot.each.dose, aes(TIME, `50%`, color=DOSE_LEVEL)) +
  facet_wrap(~DOSE_LEVEL, scale="free") +
  geom_line() +
  geom_ribbon(aes(ymin=`5%`, ymax=`95%`),alpha=0.1) +
  geom_line( data=data.pk.plot, aes(TIME,CONC,group=ID), linetype="dotted") +
  geom_point(data=data.pk.plot, aes(TIME,CONC,group=ID)) +
  ylab("CONC")
```


# Without IIV, limit Vd

```{r}
data <- 
  list(N    = nrow(data.pk),
       N_ID = nrow(data.subj),
       ID   = data.pk$ID,
       TIME = data.pk$TIME,
       DOSE = data.subj$DOSE,
       Y    = data.pk$CONC,
       VD_low = 2)

fit.stan <-
  stan(file = "mod_01_noIIV.stan", data = data)

fit.stan
```

## Prediction vs Observed

```{r ref.label="plot pred vs obs"}
```





# With IIV, limit Vd

```{r}
data <- 
  list(N    = nrow(data.pk),
       N_ID = nrow(data.subj),
       ID   = data.pk$ID,
       TIME = data.pk$TIME,
       DOSE = data.subj$DOSE,
       Y    = data.pk$CONC,
       VD_low = 2)

fit.stan <-
  stan(file = "mod_02_withIIV.stan", data = data)

fit.stan
```

## Prediction vs Observed

```{r ref.label="plot pred vs obs"}
```


## Check covariates

```{r}

fit.CLi <- 
  summary(fit.stan, pars = c("CLi"))$summary %>% 
  tbl_df() %>% 
  select(mean)
fit.VDi <- 
  summary(fit.stan, pars = c("VDi"))$summary %>% 
  tbl_df() %>% 
  select(mean)

fit.indiv.params <-
  bind_cols(tibble(ID=1:nrow(fit.CLi)),
            fit.CLi %>% rename(CL=mean),
            fit.VDi %>% rename(VD=mean))


fit.indiv.params %>% 
  ggplot(aes(CL,VD)) +
  geom_point() +
  geom_smooth(method="lm")


data.cov <- read_csv("../data/subj_dose_cov_20170521.csv") 

fit.indiv.params.cov <-
  full_join(fit.indiv.params, data.cov)

fit.indiv.params.cov %>% 
  gather(Parameter, Value, CL, VD) %>% 
  ggplot(aes(WT,Value)) +
  geom_point() +
  facet_wrap(~Parameter, scales="free") +
  geom_smooth(method="lm")

fit.indiv.params.cov %>% 
  gather(Parameter, Value, CL, VD) %>% 
  ggplot(aes(factor(SEX),Value)) +
  geom_boxplot() +
  facet_wrap(~Parameter, scales="free")


```



# With IIV, ordered

```{r}
data <- 
  list(N    = nrow(data.pk),
       N_ID = nrow(data.subj),
       ID   = data.pk$ID,
       TIME = data.pk$TIME,
       DOSE = data.subj$DOSE,
       Y    = data.pk$CONC)

fit.stan <-
  stan(file = "mod_02_withIIV_ordered.stan", data = data)

fit.stan
```


# With IIV, ordered2

```{r}
data <- 
  list(N    = nrow(data.pk),
       N_ID = nrow(data.subj),
       ID   = data.pk$ID,
       TIME = data.pk$TIME,
       DOSE = data.subj$DOSE,
       Y    = data.pk$CONC)

fit.stan <-
  stan(file = "mod_02_withIIV_ordered_2.stan", data = data)

fit.stan
```



