---
title: "Stan analysis 2 - Population PK, with or without IIV"
author: "Kenta Yoshida"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: hide
---

# Setup and load data
```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(tidyverse)
library(rstan)
library(ggmcmc)

rstan_options(auto_write=T)
options(mc.cores=parallel::detectCores())

data.pk   <- read_csv("../data/sim_pk_20170521.csv") 
data.subj <- read_csv("../data/subj_dose_20170521.csv")

data.pk.plot <- 
  mutate(data.pk, ID=factor(ID), DOSE_LEVEL=factor(DOSE_LEVEL))
```

# Summary


# Data quick look

```{r}

data.pk.plot %>% 
  ggplot(aes(TIME, CONC, group=ID, color=DOSE_LEVEL)) +
  geom_line() +
  geom_point() +
  facet_wrap(~DOSE_LEVEL) +
  scale_y_log10()

```


# Without IIV, limit Vd

```{r}
init <- function(){
    list(KA = exp(rnorm(1, log(0.5), 0.2)),
         CL = exp(rnorm(1, log(0.5), 0.2)),
         VD = exp(rnorm(1, log(5), 0.2)),
         s_Y = runif(1, 0.5, 2))
}

data <- 
  list(N    = nrow(data.pk),
       N_ID = nrow(data.subj),
       ID   = data.pk$ID,
       TIME = data.pk$TIME,
       DOSE = data.subj$DOSE,
       Y    = data.pk$CONC)

fit.stan <-
  stan(file = "mod_01_noIIV.stan", 
       data = data, init = init)

fit.stan
```

## Diagnostic plots

```{r}
# http://xavier-fim.net/packages/ggmcmc/#using-ggmcmc

fit.param <- ggs(fit.stan)
list.param <- c("KA", "CL", "VD", "s_Y")

fit.param.plot <- 
  fit.param %>% filter(Parameter %in% c(list.param))

ggs_density(fit.param.plot)
ggs_traceplot(fit.param.plot)
ggs_autocorrelation(fit.param.plot)

ggs_Rhat(fit.param.plot)
```

## Prediction vs Observed

```{r plot pred vs obs}
mcmc.sample <- rstan::extract(fit.stan)

# Plot for each subject
y.pred.interval <- 
  mcmc.sample$y_new %>% 
  apply(MARGIN=2,quantile, prob=c(0.05, 0.5, 0.95)) %>% 
  t() %>% 
  tbl_df()

bind_cols(data.pk,
          y.pred.interval) %>% 
  filter(ID<=9)  %>% 
  mutate(ID=factor(ID)) %>% 
  ggplot(aes(TIME, `50%`, color=ID)) +
  facet_wrap(~ID) +
  geom_line() +
  geom_ribbon(aes(ymin=`5%`, ymax=`95%`, fill=ID),alpha=0.1) +
  geom_point(data=data.pk %>% filter(ID<=9) %>% mutate(ID=factor(ID)),
             aes(TIME,CONC)) +
  ylab("CONC")


# Plot for each dose level

dat.fit.plot.each.dose <- 
  ## Convert to tibble
  mcmc.sample$y_new %>% t() %>% tbl_df() %>% 
  ## Re-groupe predicted values according to TIME and DOSE_LEVEL
  bind_cols(data.pk,.) %>% 
  gather(key,prediction,-(ID:DOSE_LEVEL)) %>% 
  group_by(TIME,DOSE_LEVEL) %>%
  ## Calc quantiles
  summarize(`5%` =quantile(prediction, probs=0.05),
            `50%`=quantile(prediction, probs=0.5),
            `95%`=quantile(prediction, probs=0.95)) %>% 
  ungroup() %>% 
  mutate(DOSE_LEVEL=factor(DOSE_LEVEL))

ggplot(dat.fit.plot.each.dose, aes(TIME, `50%`, color=DOSE_LEVEL)) +
  facet_wrap(~DOSE_LEVEL, scale="free") +
  geom_line() +
  geom_ribbon(aes(ymin=`5%`, ymax=`95%`),alpha=0.1) +
  geom_line( data=data.pk.plot, aes(TIME,CONC,group=ID), linetype="dotted") +
  geom_point(data=data.pk.plot, aes(TIME,CONC,group=ID)) +
  ylab("CONC")
```


# With IIV

```{r}
init <- function(){
    list(KA = exp(rnorm(1, log(0.5), 0.2)),
         CL = exp(rnorm(1, log(0.5), 0.2)),
         VD = exp(rnorm(1, log(5), 0.2)),
         s_CL= exp(rnorm(1, log(0.2), 0.5)),
         s_VD= exp(rnorm(1, log(0.2), 0.5)),
         s_Y = runif(1, 0.5, 2),
         CLi = rep(0.5,nrow(data.subj)),
         VDi = rep(5,  nrow(data.subj)))
}

data <- 
  list(N    = nrow(data.pk),
       N_ID = nrow(data.subj),
       ID   = data.pk$ID,
       TIME = data.pk$TIME,
       DOSE = data.subj$DOSE,
       Y    = data.pk$CONC)

fit.stan <-
  stan(file = "mod_02_withIIV.stan", 
       data = data, init = init)

fit.stan
```

## Diagnostic plots

```{r}
# http://xavier-fim.net/packages/ggmcmc/#using-ggmcmc

fit.param <- ggs(fit.stan)
list.param.1 <- c("KA", "CL", "VD")
list.param.2 <- c("s_CL", "s_VD", "s_Y")

fit.param.plot <- 
  fit.param %>% filter(Parameter %in% c(list.param.1,list.param.2))
fit.param.plot.1 <- 
  fit.param %>% filter(Parameter %in% list.param.1)
fit.param.plot.2 <- 
  fit.param %>% filter(Parameter %in% list.param.2)

ggs_density(fit.param.plot.1)
ggs_density(fit.param.plot.2)
ggs_traceplot(fit.param.plot.1)
ggs_traceplot(fit.param.plot.2)
ggs_autocorrelation(fit.param.plot.1)
ggs_autocorrelation(fit.param.plot.2)

ggs_Rhat(fit.param.plot)
```

## Prediction vs Observed
### Individual parameter estimates (IPRED)

```{r ref.label="plot pred vs obs"}
```

### Population parameter estimates + IIV

```{r}
mcmc.sample <- rstan::extract(fit.stan)

# Plot for each subject
y.pred.interval <- 
  mcmc.sample$y_newPred %>% 
  apply(MARGIN=2,quantile, prob=c(0.05, 0.5, 0.95)) %>% 
  t() %>% 
  tbl_df()

bind_cols(data.pk,
          y.pred.interval) %>% 
  filter(ID<=9)  %>% 
  mutate(ID=factor(ID)) %>% 
  ggplot(aes(TIME, `50%`, color=ID)) +
  facet_wrap(~ID) +
  geom_line() +
  geom_ribbon(aes(ymin=`5%`, ymax=`95%`, fill=ID),alpha=0.1) +
  geom_point(data=data.pk %>% filter(ID<=9) %>% mutate(ID=factor(ID)),
             aes(TIME,CONC)) +
  ylab("CONC")


# Plot for each dose level

dat.fit.plot.each.dose <- 
  ## Convert to tibble
  mcmc.sample$y_newPred %>% t() %>% tbl_df() %>% 
  ## Re-groupe predicted values according to TIME and DOSE_LEVEL
  bind_cols(data.pk,.) %>% 
  gather(key,prediction,-(ID:DOSE_LEVEL)) %>% 
  group_by(TIME,DOSE_LEVEL) %>%
  ## Calc quantiles
  summarize(`5%` =quantile(prediction, probs=0.05),
            `50%`=quantile(prediction, probs=0.5),
            `95%`=quantile(prediction, probs=0.95)) %>% 
  ungroup() %>% 
  mutate(DOSE_LEVEL=factor(DOSE_LEVEL))

ggplot(dat.fit.plot.each.dose, aes(TIME, `50%`, color=DOSE_LEVEL)) +
  facet_wrap(~DOSE_LEVEL, scale="free") +
  geom_line() +
  geom_ribbon(aes(ymin=`5%`, ymax=`95%`),alpha=0.1) +
  geom_line( data=data.pk.plot, aes(TIME,CONC,group=ID), linetype="dotted") +
  geom_point(data=data.pk.plot, aes(TIME,CONC,group=ID)) +
  ylab("CONC")
```

## Parameter distributions
### Individual parameter estimates (IPRED)

```{r}
fit.CLi <- 
  summary(fit.stan, pars = c("CLi"))$summary %>% 
  tbl_df() %>% 
  select(mean)
fit.VDi <- 
  summary(fit.stan, pars = c("VDi"))$summary %>% 
  tbl_df() %>% 
  select(mean)

fit.indiv.params <-
  bind_cols(tibble(ID=1:nrow(fit.CLi)),
            fit.CLi %>% rename(CL=mean),
            fit.VDi %>% rename(VD=mean))

fit.indiv.params %>% 
  ggplot(aes(CL,VD)) +
  geom_point() +
  geom_smooth(method="lm")
```

### Population parameter estimates + IIV

```{r}
mcmc.sample <- rstan::extract(fit.stan)

fit.CLiPred <- mcmc.sample$CLiPred[1,] 
fit.VDiPred <- mcmc.sample$VDiPred[1,] 


fit.indiv.params.pred <-
  tibble(CL = fit.CLiPred,
         VD = fit.VDiPred)

fit.indiv.params.pred %>% 
  ggplot(aes(CL,VD)) +
  geom_point() +
  geom_smooth(method="lm")
```

### Combined plot

```{r}
bind_rows(fit.indiv.params %>% mutate(TYPE="Estimates"),
          fit.indiv.params.pred %>% mutate(TYPE="Predicted")) %>% 
  ggplot(aes(CL,VD,color=TYPE, fill=TYPE)) +
  geom_point() +
  geom_smooth(method="lm", alpha=0.1)
```

## Check covariates

```{r}

data.cov <- read_csv("../data/subj_dose_cov_20170521.csv") 

fit.indiv.params.cov <-
  full_join(fit.indiv.params, data.cov)

fit.indiv.params.cov %>% 
  gather(Parameter, Value, CL, VD) %>% 
  ggplot(aes(WT,Value)) +
  geom_point() +
  facet_wrap(~Parameter, scales="free") +
  geom_smooth(method="lm")

fit.indiv.params.cov %>% 
  gather(Parameter, Value, CL, VD) %>% 
  ggplot(aes(factor(SEX),Value)) +
  geom_boxplot() +
  facet_wrap(~Parameter, scales="free")


```
